---
title: "ML Final Project"
output:
  word_document: default
  pdf_document: default
---


# Loading the necessary Libraries for project

```{r}
library(dplyr)
library(caret)
library(missForest)
library(corrplot)
library(factoextra)
library(cluster)
```

# Reading the CSV File

```{r}
# reading file
Fuel_Receipts_Costs_Data=read.csv("C:/Users/Pavan Chaitanya/Downloads/fuel_receipts_costs_eia923 (1).csv")

# head part of file
head(Fuel_Receipts_Costs_Data,5)

#Checiking NA's
colMeans(is.na(Fuel_Receipts_Costs_Data))

```

# Data Cleaning and Removing the Unnecessary Colums that are present in dataset

```{r}
# Randonmly Assigning the seed value
set.seed(2875)

#checking the NA Values
Fuel_Receipts_Costs_Data[Fuel_Receipts_Costs_Data==""] = NA

#Converting the mean values to the percentage
Filtering_NA  = Fuel_Receipts_Costs_Data[,(colMeans(is.na(Fuel_Receipts_Costs_Data))*100)<50]

#Sampling the 2 % of the data 
Creating_Two_data_Partition = createDataPartition(Filtering_NA$plant_id_eia,p=0.02,list = FALSE)
Creating_Two_data_Partition1 = Filtering_NA[Creating_Two_data_Partition,]
# Printing the 2% data
head(Creating_Two_data_Partition1,10)

colMeans(is.na(Creating_Two_data_Partition1))*100

#converting the date to date format
Creating_Two_data_Partition1$report_date <- as.Date(Creating_Two_data_Partition1$report_date)

Creating_Two_data_Partition1$report_date <- as.numeric(format(Creating_Two_data_Partition1$report_date, "%Y"))

# removing the unnecessary Colums
Creating_Two_data_Partition1=Creating_Two_data_Partition1[,-c(6,8,17)]

# Printing the data data frame after removing unnecessary columns
head(Creating_Two_data_Partition1,10)

```

# Data Imputation

```{r}
# Converting the variables of char to factor type for data impuataion 
Creating_Two_data_Partition1$report_date = as.factor(Creating_Two_data_Partition1$report_date)

Creating_Two_data_Partition1$contract_type_code = as.factor(Creating_Two_data_Partition1$contract_type_code)

Creating_Two_data_Partition1$energy_source_code = as.factor(Creating_Two_data_Partition1$energy_source_code)

Creating_Two_data_Partition1$fuel_group_code = as.factor(Creating_Two_data_Partition1$fuel_group_code)

Creating_Two_data_Partition1$primary_transportation_mode_code = as.factor(Creating_Two_data_Partition1$primary_transportation_mode_code)

Creating_Two_data_Partition1$natural_gas_transport_code = as.factor(Creating_Two_data_Partition1$natural_gas_transport_code)

# Computing the Data Imputation
Genertated_Data = missForest(Creating_Two_data_Partition1)

#Taking only the ximp data frame 
Imputed = Genertated_Data$ximp

#Printing the data frame after computation of the missing values
head(Imputed,10)



```

# Partitioning the  2 % data into 75 % training data.  
```{r}

Data_Partition = createDataPartition(Imputed$plant_id_eia,p=0.75,list = FALSE)

Data_Partition_Trained = Imputed[Data_Partition,]

Data_Partition_Tested = Imputed[-Data_Partition,]

```

# As data has Outliers we are making sure that the outlier are removed.

```{r}
# For the fuel received units performing the quartile ranges and IQR  
Quartiled_data = quantile(Data_Partition_Trained$fuel_received_units, probs=c(.25, .75), na.rm = FALSE)
Data_Partition_Quartiled = IQR(Data_Partition_Trained$fuel_received_units)


Fuelunits_Lower = Quartiled_data[1] - 1.5*Data_Partition_Quartiled
Fuelunits_Upper = Quartiled_data[2] + 1.5*Data_Partition_Quartiled 
 
Data_With_No_Outliers = subset(Data_Partition_Trained, Data_Partition_Trained$fuel_received_units > Fuelunits_Lower & Data_Partition_Trained$fuel_received_units < Fuelunits_Upper)

# For the fuel cost per mmbtu performing the quartile ranges and IQR
Range_of_Fuel = quantile(Data_With_No_Outliers$fuel_cost_per_mmbtu, probs=c(.25, .75), na.rm = FALSE)
Fuelcost_IQR <- IQR(Data_With_No_Outliers$fuel_cost_per_mmbtu)
 
Fuelcost_Lower = Range_of_Fuel[1] - 1.5*Fuelcost_IQR
Fuelcost_Upper = Range_of_Fuel[2] + 1.5*Fuelcost_IQR 
 
No_Outlier_Data = subset(Data_With_No_Outliers, Data_With_No_Outliers$fuel_cost_per_mmbtu > Fuelcost_Lower & Data_With_No_Outliers$fuel_cost_per_mmbtu < Fuelcost_Upper)

```
# Choosing and Normalising the  selected variables

```{r}

All_Numeric_Variables=No_Outlier_Data[,c(7,8,9,10,11,12)]
head(All_Numeric_Variables,12)

Scaled_Data = scale(All_Numeric_Variables)
head(Scaled_Data,12)

```

# K-Means Clustering 

```{r}
#wss
fviz_nbclust(Scaled_Data, kmeans, method = "wss")
# We feel that k=2 is best.
wss_k2 = kmeans(Scaled_Data, centers=2,nstart=50)
wss_group=wss_k2$cluster
wss_k2$withinss
wss_k2$tot.withinss


fviz_nbclust(Scaled_Data, kmeans, method = "silhouette")
# Silhouette shows that k=3 is best.
Sil_k3 = kmeans(Scaled_Data, centers=3,nstart=50)
Silhouette_group=Sil_k3$cluster
Sil_k3$withinss
Sil_k3$tot.withinss

# By comparing the both methods  and by finding the withiness we have come to an idea that k=3 is the best k for our project. 
# ie Sil_k3$tot.withinss is less that of Wss_k2$tot.withinss
# 2101.58 is less than 25135.54

fviz_cluster(Sil_k3,data=Scaled_Data)

```
Interpretation
```{r}
Silhouette_group = as.data.frame(Silhouette_group)
Sil_bind=cbind(All_Numeric_Variables,Silhouette_group)
Cluster_mean= Sil_bind %>% group_by(Silhouette_group) %>%
summarise_all("mean")
Cluster_mean

# 
# As Sulfer content,ash content,mercury content are less than 0.001 m they can be neglected for intrepretation.
# Cluster  1
# 
#  The power plants present in this cluster recieves fuel of 18049.93 which is low compared to other plants.
#  As they are receiving low fuel their heat content in fuel(fuel_mmbtu) is also low (1.183889).
#  The fuel cost per mmbtu is higher(4.889421) than all the 3 clusters formed.
#  This Cluster is not a preferred one to recommend for Us Government because of fuel cost per mmbtu.
# 
# Cluster 2
#  
#   The power Plants present in this cluster receives fuel of 29500.21 which is slightly above the Cluster 3 but not cluster 1.
#   Their heat content in the fuel is very very high of 21.607668 comapared to all the 3 clsuters.
#   The fuel cost per mmbtu is lower(2.635552) than all the 3 clusters formed.
#   This cluster is also not a preferred one to recommend for us Government because of fuel mmbtu per unit.
#   
# Cluster 3
#  
# 	The power Plants present in this cluster receives fuel of 161115.82 which is high than all the 3 clsuters.
# 	Their heat content in the fuel is 5.231477 which is very good wrt to the fuel recieves compared to other 2 clsuters.
# 	The fuel cost per mmbtu is also very good(3.704139) wrt to fuel recieved and the heat content.
# 	This Cluster is the preferred one to recommend for the Us Government beacuse by looking all the factors like (fuel recieved,heat content,fuel cost per mmbtu).
```


# Hierarchial Clustering for visualizing the data 
 
```{r}
# Getting distance
distance= dist(Scaled_Data,method="euclidean")
# Computing method
hclust_ward=hclust(distance,method = "ward.D2")
#plotting 
plot(hclust_ward,cex=0.6,hang=-1); 
rect.hclust(hclust_ward,k=3,border=1:4)

```






