---
title: "Assignment_1"
author: "Pavan Chaitanya"
date: "9/9/2022"
output: word_document
 
---

#ImportData 
```{r}
library(readr)
food_coded <- read_csv("C:\\Users\\Pavan Chaitanya\\Documents\\ML\\food_coded.csv")
```
# viewing the imported file in the R studio
```{r}
View(food_coded)
```
# printing the descriptive Statistics for quantitative and categorical variables
 # For quantitative variables 
  # 1 min function
```{r}     
min(food_coded$calories_chicken)
```  
  # 2 max function
```{r}     
max(food_coded$calories_chicken)
```        
  # 3 Mean Function
```{r}     
mean(food_coded$indian_food)
```    
 # 4 Median Function
```{r}     
median(food_coded$ideal_diet_coded)
 
```  
 # 5 Mode Function
       # there is no direct function available for mean. We have to group first then perform mean.
```{r}      
   food<-table(food_coded$vitamins)
        sort(food,decreasing = TRUE ) 

```  
 # For categorial variables
     # 1 Range Function     
```{r}     
        range(food_coded$comfort_food)
        range(food_coded$food_childhood)
```  
     # 2 Summary Function 
```{r}     
        summary(food_coded)
```  


# Tranformation of variables.
  # printing the original Healthy feeling Values      
```{r}     
      print(food_coded$healthy_feeling)
```  
  # Applying the tranformation(logrithmic)
```{r}     
      healthyfealinglog<-log10(food_coded$healthy_feeling)
```  
  # printing the Logrithmic Healthy feeling Values         
```{r}     
      print(healthyfealinglog)
```  
  # printing the original ethnic food Values   
```{r}     
      print(food_coded$ethnic_food)
```  
  # Applying the tranformation(sqrt)      
```{r}     
      sqrt_ethnicfood<-sqrt(food_coded$ethnic_food)
```  
  # printing the squared ethnic food Values  
```{r}     
    print(sqrt_ethnicfood)
```  


#plotting one quantitave Variable
  #We will be using histogram to plot the quantitative Variable
    # plotting the histogram for healthy feeling variable 
```{r}     
      hist(food_coded$healthy_feeling, col = 'red',main = 'Original Histogram')
```  
    # plotting the histogram for income variable 
```{r}     
      hist(food_coded$income, col = 'yellow', main = 'income')
```  


#plotting the Scatter plot
  #plotting the scatter plot between food_reasons_coded...12 and food_reasons_coded...10
```{r}     
      plot(food_coded$comfort_food_reasons_coded...12,food_coded$comfort_food_reasons_coded...10,main = 'Plotting Scatter Plot between the 2 comfort food resons',xlab = 'Food Resaon 12', ylab ='Food Reason 10' )
```  


